### 2023-09-19 ms/test.Rmd

---
title: "test"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dispersion, fig.cap = "Stomata are more dispersed than expected under the null model of uniform random position (dispersion index = 0) but far from a distribution that maximizes distance between stomata (dispersion index = 1)", echo = TRUE}

knitr::include_graphics("figures/dispersion.pdf")

```