### 2023-09-25 ms/si.qmd
starting model description in SI

---
title: "2-D Photosynthesis Model"
format: docx
editor: visual
bibliography: ../references.bib
---

We model leaf photosynthesis using a two-dimensional porous medium approximation. The model is solved using a finite element method (FEM) in the *R* package **deSolve** [@soetaert2010]. Table 1 is a glossary model terms and symbols. Here we describe the model and associated *R* code.

## Leaf anatomy

We assume that the leaf is a homogenous 2-D medium. In the final version, we will incorporate differences in spongy and palisade porosity, gradients in light absorption, electron transport capacity, and Rubisco concentration. The mesophyll is $T_\text{leaf}$ thick and the stomata are regularly spaced apart by distance $U$ on both ab- and adaxial surfaces. In this scenario, we assume that the stomata on each surface are precisely offset from each other by distance $U/2$. This minimizes the average distance between any point in the mesophyll and its nearest stomate. Because of the regular spacing, we only need to model the region between a stomate on surface and the next stomate on the other surface (fig. 1). The rest of the mesophyll will be the same because of symmetry.

```{r, message = FALSE, fig.cap='Figure 1: Example leaf anatomy analyzed by the 2-D FEM.', echo = FALSE, fig.height=6, fig.width=3}

library(dplyr)
library(ggforce)
library(ggplot2)
library(tidyr)

# Example leaf where leaf thickness equals interstomatal distance
 xz_ratio = 0.5

  df_stomata = crossing(
    nesting(x = c(0, 1), y = c(0, 1) / xz_ratio),
    r = 0.04,
    x_offset = c(-1.5, 1.5)
  ) |>
    mutate(x0 = x + x_offset * r)

  df_epidermis = df_stomata |>
    summarise(stomata_min = min(x0 - r), stomata_max = max(x0 + r), .by = "y") |>
    mutate(
      xmin = ifelse(stomata_min < 0, stomata_max, 0),
      xmax = ifelse(stomata_max > 1, stomata_min, 1)
    )

  ggplot() +
    geom_circle(data = df_stomata, mapping = aes(x0 = x0, y0 = y, r = r)) +
    geom_segment(data = df_epidermis,
                 mapping = aes(
                   x = xmin,
                   xend = xmax,
                   y = y,
                   yend = y
                 ), linewidth = 1.5) +
    geom_text(
      data = df_stomata |>
        select(x, y) |>
        distinct(),
      mapping = aes(x, y, label = "stomate"),
      position = position_nudge(y = c(-0.05, 0.05) / xz_ratio)
    ) +
    geom_text(
      mapping = aes(x = 0.5, y = c(-0.1, 0.5, 1.1) / xz_ratio, label = c("abaxial", "mesophyll", "adaxial"))
    ) +
    geom_segment(mapping = aes(x = 0, y = -0.15 / xz_ratio, xend = 1, yend = -0.15 / xz_ratio),
                 arrow = arrow(angle = 90, length = unit(0.0125 / xz_ratio, "npc"), ends = "both") ) +
    geom_segment(mapping = aes(x = -0.15, y = 0 / xz_ratio, xend = -0.15, yend = 1 / xz_ratio),
                 arrow = arrow(angle = 90, length = unit(0.0125 / xz_ratio, "npc"), ends = "both") ) +
    geom_text(mapping = aes(x = 0.5, y = -0.2 / xz_ratio, label = "paste(italic(U) / 2, ', half the interstomatal distance [', mu, 'm]')"),
              parse = TRUE) +
    geom_text(mapping = aes(x = -0.2, y = 0.5 / xz_ratio, label = "paste(italic(T)[leaf], ', leaf thickness [', mu, 'm]')"),
              parse = TRUE, angle = 90) +
    coord_equal(clip = "off") +
    theme_void()

```