### 2023-10-16 ms/test.Rmd
added model_var_table

---
title: "test"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
knitr::opts_chunk$set(out.width = '100%', dpi = 300) # Figure resolution and size
knitr::opts_chunk$set(fig.env = "figure") # Latex figure environment
knitr::opts_knit$set(root.dir = "..")
```

```{r}
source("r/header.R")
fit = readr::read_rds("objects/fit.rds")
model_var_table = readr::read_rds("objects/model_var_table.rds")
```

```{r, model_var}
knitr::kable(model_var_table)
```