### 2023-10-17 ms/ms.Rmd
added correlation_anova to objects

---
# This template is largely functionally equivalent to that for elsevier
# The two fields that are not in the elsevier format are:
#      `corresponding_author` and `acknowledgements`
title: Spatial Coordination Of Stomatal Patterning Between Leaf Surfaces In Amphistomatous *Arabidopsis thaliana* Incurs No Photosynthetic Advantage
#date: "`r Sys.Date()`"
author:
  - name: Jacob L. Watts
    email: jwatts@colgate.edu
    corresponding_author: yes # Footnote using corresponding_author
    affiliation: Colgate University
  - name: Graham Dow
    email: graham.dow@usys.ethz.ch
    affiliation: ETH
  - name: Thomas N. Buckley
    email: tnbuckley@ucdavis.edu
    affiliation: University of California Davis
    footnote: 1 # Footnote using reference
  - name: Christopher D. Muir
    email: cdmuir@hawaii.edu
    affiliation: University of Hawaii at Manoa
    footnote: 1 # Footnote using reference
address:
  - code: Colgate University
    address: 13 Oak Drive, Hamilton, NY 13346
  - code: University of Hawaii at Manoa
    address: 2500 Campus Rd, Honolulu, HI 96822
footnote:
  - code: 1
    text: "Equal contribution"
  - code: 2
    text: "Current email address: \\href{mailto:cat@example.com}{cat@example.com}"
abstract: |
  This is the abstract.

  It consists of two paragraphs.
acknowledgements: |
  This is an acknowledgement.

  It consists of two paragraphs.
keywords:
  - amphistomy
  - Arabidopsis thaliana
  - CO$_2$ diffusion
  - finite element method
  - optimality
  - photosynthesis
  - stomata
fontsize: 12pt
#spacing: halfline # could also be oneline
#classoptions:
#  - endnotes
bibliography: stomata-spacing.bib
#csl: mycitationstyle.csl
link-citations: yes
output: 
  rticles::oup_article:
    latex_engine: xelatex
#urlcolor: orange
#linkcolor: green
#citecolor: red
header-includes:
 - \usepackage[nomarkers,tablesfirst]{endfloat} # For figures and tables at end
 - \usepackage{lineno} # For line numbering
 - \linenumbers # For line numbering
 - \usepackage{hyperref} # For autoref
 - \renewcommand{\figureautorefname}{Fig.}
 - \usepackage{booktabs} # for tables
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
knitr::opts_chunk$set(out.width = '100%', dpi = 300) # Figure resolution and size
knitr::opts_chunk$set(fig.env = "figure") # Latex figure environment
knitr::opts_knit$set(root.dir = "..")
```

```{r}
source("r/header.R")
fit = readr::read_rds("objects/fit.rds")
single_surface_results = readr::read_rds("objects/single_surface_results.rds")
single_surface_anova = readr::read_rds("objects/single_surface_anova.rds")
density_anova = readr::read_rds("objects/density_anova.rds")
dual_surface_results = readr::read_rds("objects/dual_surface_results.rds")
correlation_anova = readr::read_rds("objects/correlation_anova.rds")
pm_pars = readr::read_csv("raw-data/2d-pm-parameters.csv", show_col_types = FALSE)
model_var = readr::read_rds("objects/model_var.rds")

# Summarize results on NNI
nni1 = single_surface_results |>
  dplyr::transmute(sig = p_value < 0.05) |>
  dplyr::summarize(n1 = sum(sig), n2 = sum(!sig))

adjp = mt.rawp2adjp(single_surface_results$p_value, proc = "BH")

nni2 = bind_cols(single_surface_results[adjp$index,], adjp$adjp) |>
  dplyr::transmute(sig = BH < 0.05) |>
  dplyr::summarize(n1 = sum(sig), n2 = sum(!sig))

# Summarize results on corr(NSD)
nsd1 = dual_surface_results |>
  ungroup() |>
  dplyr::transmute(sig = p_value < 0.05) |>
  dplyr::summarize(n1 = sum(sig), n2 = sum(!sig))

adjp = mt.rawp2adjp(dual_surface_results$p_value, proc = "BH")

nsd2 = bind_cols(dual_surface_results[adjp$index,], adjp$adjp) |>
  ungroup() |>
  dplyr::transmute(sig = BH < 0.05) |>
  dplyr::summarize(n1 = sum(sig), n2 = sum(!sig))

```

# Introduction

Stomatal anatomy (e.g. size, density, distribution, and patterning) and movement regulate gas exchange during photosynthesis, namely CO$_2$ assimilation and water loss through transpiration. Since waxy cuticles are mostly impermeable to CO$_2$ and H$_2$O, stomata are the primary entry points through which gas exchange occurs despite making up a small percentage of the leaf area [@lange_responses_1971]. Stomata consist of two guard cells which open and close upon changes in turgor pressure or hormonal cues [@mcadam_linking_2016]. The stomatal pore leads to an internal space known as the substomatal cavity where gases contact the mesophyll. Once in the mesophyll, CO$_2$ diffuses throughout a network of intercellular air space (IAS) and into mesophyll cells where CO$_2$ assimilation ($A$) occurs within the chloroplasts [@lee_diffusion_1964]. Stomatal conductance and transpiration are determined by numerous environmental and anatomical parameters such as vapor pressure deficit (VPD), irradiance, temperature, wind speed, leaf water potential, IAS geometry, mesophyll cell anatomy, and stomatal anatomy.

Many successful predictions about stomata and other leaf traits can be made by hypothesizing that natural selection should optimizes CO$_2$ gain per unit of water loss [@cowan_stomatal_1977; @buckley_optimal_2017; @sperry_predicting_2017]. However, stomatal anatomy may be partially constrained by physical and developments limits on phenotypic expression [@croxdale_stomatal_2000; @harrison_influence_2020; @muir_how_2023]. Sometimes optimization leads to similar phenotypes across many disparate species. For example, almost all stomata follow the one cell spacing rule to maintain proper stomatal functioning [@geisler_oriented_2000; @dow_physiological_2014]; however some species (notably in *Begonia*) appear to benefit from overlapping vapor shells caused by stomatal clustering [@yi_gan_stomatal_2010; @lehmann_effects_2015; @papanatsiou_stomatal_2017]. Stomatal traits also vary adaptively in different environments. Stomatal density positively co-varies with irradiance during leaf development and negatively co-varies with CO$_2$ concentration [@gay_influence_1975; @schoch_dependence_1980; @woodward_stomatal_1987; @royer_stomatal_2001], consistent with optimality predictions. Stomatal size is jointly controlled by genome size, light, and stomatal density [@jordan_environmental_2015]. Size positively co-varies with genome size [@roddy_scaling_2020] and negatively co-varies with stomatal density [@camargo_density_2011]. Total stomatal area (size $\times$ density) is optimized for operational conductance ($gs_\text{s,op}$) rather than maximum conductance ($gs_\text{s,max}$) such that stomatal apertures are most responsive to changes in the environment at their operational aperture [@franks_physiological_2012; @liu_scaling_2021]. Stomatal aperture can compensate for maladaptive stomatal densities to an extent [@bussis_stomatal_2006], but stomatal density and size ultimately determine a leaf's theoretical $gs_\text{s,max}$ [@sack_developmental_2016], which is proportional to $gs_\text{s,op}$ [@murray_consistent_2020]. Additionally, low stomatal densities lead to irregular and insufficient CO$_2$ supply and reduced photosynthetic efficiency in areas far from stomata [@pieruschka_lateral_2006; @morison_lateral_2005], while high stomatal densities can reduce water use efficiency (WUE) [@bussis_stomatal_2006] and incur excessive metabolic costs [@deans_optimization_2020]. In most species, stomata occur on the abaxial (usually lower) leaf surface; but amphistomy, the occurrence of stomata on both abaxial and adaxial leaf surfaces, is also prevalent in high light environments with constant or intermittent access to sufficient water [@mott_adaptive_1982; @jordan_using_2014; @muir_light_2018; @drake_two_2019; @muir_is_2019]. Amphistomy effectively halves the CO$_2$ diffusion path length and boundary layer resistance by doubling boundary layer conductance [@parkhurst_adaptive_1978; @harrison_influence_2020; @mott_amphistomy_1991]. Historically, stomatal patterning in dicot angiosperms was thought to be random with an exclusionary distance surrounding each stomate [@sachs_developmental_1974]; however, the developmental controls of stomatal patterning are poorly understood and likely more complex than random development along the leaf surface. @croxdale_stomatal_2000] reviews three developmental theories which attempt to explain stomatal patterning in angiosperms: inhibition, cell lineage, and cell cycle, ultimately arguing for a cell cycle based control of stomatal patterning.

The patterning and spacing of stomata on the leaf affects photosynthesis in $C_3$ leaves by altering the CO$_2$ diffusion path length from stomata to sites of carboxylation in the mesophyll. Maximum photosynthetic rate ($A_\text{max}$) in $C_3$ plants is generally co-limited by biochemistry and diffusion, but modulated by light availability [@parkhurst_intercellular_1990; @manter_ci_2004; @carriqui_diffusional_2015]. Low light decreasing CO$_2$ demand by limiting electron transport rate, leading to relatively high internal CO$_2$ concentration ($C_\text{i}$) and low $A_\text{max}$ [@kaiser_metabolic_2016]. In contrast, well hydrated leaves with open stomata in high light, photosynthesis is often limited by CO$_2$ supply as resistances from the boundary layer, stomatal pore, and mesophyll can result in insufficient CO$C_2$ supply at the chloroplast to maxmimize photosynthesis [@farquhar_biochemical_1980; @lehmeier_cell_2017]. In this study, we focus primarily on how stomatal patterning affects diffusion, ignoring boundary layer and mesophyll resistances.

To maximize CO$_2$ supply from the stomatal pore to chloroplasts, stomata should be uniformly distributed in an equilateral triangular grid on the leaf surface so as to minimize stomatal number and CO$_2$ diffusion path length [@parkhurst_diffusion_1994]. As the diffusion rate of CO$_2$ though liquid is approximately $10^4\times$ slower than CO$_2$ diffusion through air, mesophyll resistance is generally thought to be primarily limited by liquid diffusion [@aalto_three-dimensional_2002; @evans_resistances_2009], but diffusion through the IAS has also been shown to be a rate limiting process because the tortuous, disjunct nature of the IAS can greatly increase diffusion path lengths [@harwood_understanding_2021]. Additionally, tortuosity is higher in horizontal directions (parallel to leaf surface) than vertical directions (perpendicular to leaf surface) because of the cylindrical shape and vertical arrangement of pallisade mesophyll cells [@earles_beyond_2018; @harwood_understanding_2021]. However, the ratio of lateral to vertical diffusion rate is still largely unknown [@morison_lateral_2005; @pieruschka_lateral_2005; @pieruschka_lateral_2006]. Depending on the thickness of the leaf, porosity of the leaf mesophyll, tortuosity of the IAS, and lateral to vertical diffusion rate ratio, minimizing diffusion path length for CO$_2$ via optimally distributed stomata may yield significant increases in CO$_2$ supply for photosynthesis and higher $A_\text{max}$.

We hypothesized that natural selection will favor stomatal patterning and distribution to minimize the diffusion path length. In amphistomatous leaves, this would be accomplished by 1) a uniform distribution of stomata on both abaxial and adaxial leaf surfaces and 2) coordinated stomatal spacing on each surface that offsets the position of stomata (\autoref{fig:ideal-amphi-grid}). Coordination between leaf surfaces is defined in this study as the occurrence of stomata in areas farthest from stomata on the opposite leaf surface. Additionally, because CO$_2$ is more limiting for photosynthesis under high light, we hypothesize that in high light 3) there should be more stomata, and 4) stomata should be more uniformly distributed than in low light. Finally, as stomatal densities are selected for optimal operational aperture, we hypothesize that 5) stomatal length will be positively correlated with the area of the leaf surface to which it is closest. We refer to this as the 'stomatal zone', the leaf area surrounding a focal stomate closest to that stomate and therefore the zone it supplies with CO$_2$). This way, each stomate can be optimally sized relative to the mesophyll volume it supplies.

\begin{figure}[ht]
\includegraphics[width = \textwidth]{figures/ideal-amphi-grid.pdf}
\caption{Idealized amphistomatous stomatal grid with uniform stomatal patterning and perfect abaxial-adaxial coordination.}
\label{fig:ideal-amphi-grid}
\end{figure}

To test these hypotheses, we grew the model plant *Arabidopsis thaliana* in high, medium, and low light and measured stomatal density, size, and patterning on both leaf surfaces, and spatial coordination between them. We use Voronoi tessellation techniques to calculate stomatal zones. We also used a 2-D porous medium approximation of CO$_2$ diffusion and photosynthesis to predict the photosynthetic advantage of optimal versus suboptimal coordination in stomatal coordination between surfaces. Specifically, we predicted that traits which affect diffusion path length (leaf thickness, stomatal density, leaf porosity, lateral-vertical diffusion rate ratio), diffusion rate (temperature, pressure), and CO$_2$ demand (Rubisco concentration, light) would modulate the advantage of optimal stomatal arrangement following the relationships outlined in \autoref{tab:hypotheses}. Here, we integrate over reasonable parameter space to determine the ecophysiological context most likely to favor stomatal spatial coordination in amphistomatous leaves.

```{r hypotheses, results="asis"}

df = tibble::tribble(
  ~trait, ~relationship,
  "leaf thickness", "+",
  "stomatal density", "-",
  "leaf porosity", "-", 
  "lat.-vert. diffusion ratio", "-", 
  "temperature", "-",
  "pressure", "-",
  "Rubisco concentration", "+",
  "light", "+"
)

# Creates tables that follow OUP guidelines using xtable
print(xtable::xtable(df, caption = "A summary of the hypothesized relationships between leaf traits and environmental conditions and photosynthetic advantage of stomatal spatial coordination in amphistomatous leaves.", label = "tab:hypotheses"),
      comment = FALSE, include.rownames = FALSE)
```

# Materials and methods

## Data Preparation

[CDM: Graham Dow provided these images. We'll need to add him as a co-author and ask him to write methods on image acquisition.]

*Arabidopsis thaliana* plants were grown in three different light environments: low light (PAR = 50 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), medium light (100 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), and high light (200 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$). PAR stands for photosynthetically active radiation. Once leaves were mature, we captured images of the abaxial and adaxial leaf surfaces using XXX. We captured 132 images in total, making 66 abaxial-adaxial image pairs. We measured stomatal position and size using ImageJ [@schneider_nih_2012].

## Single surface analyses

We tested whether stomata are non-randomly distributed by comparing the observed stomatal patterning to a random uniform pattern. For each leaf surface image with $n$ stomata we generated $10^3$ synthetic surfaces with $n$ stomata uniformly randomly distributed on the surface. For each sample image, we compared the observed Nearest Neighbor Index ($\mathrm{NNI}$) to the null distribution of $\mathrm{NNI}$ values calculated from the synthetic data set. $\mathrm{NNI}$ is the ratio of observed mean distance ($\overline{D}_O$) to the expected mean distance ($\overline{D}_E$) where $\overline{D}_E$ is:

\begin{equation}\label{eq:emd}
  \overline{D}_E = \frac{0.5}{\sqrt{A_\text{leaf} / n_\text{stomata}}}.
\end{equation}

\noindent $A_\text{leaf}$ is leaf area visible in the sampled field and $n_\text{stomata}$ the number of stomata. $\overline{D}_E$ is the theoretical average distance to the nearest neighbor of each stomate if stomata were uniformly randomly distributed [@clark_distance_1954]. $\overline{D}_O$ calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \overline{D}_O = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}},
\end{equation}

\noindent where $d_i$ is the distance between $\text{stomate}_i$ and its nearest neighbor. We calculated $\mathrm{NNI}$ using the *R* package **spatialEco** version `r packageVersion("spatialEco")` [@evans_spatialeco_2023]. The observed stomatal distribution is dispersed relative to a uniform random distribution if the observed $NNI$ is greater than 95% of the synthetic $\mathrm{NNI}$ values (one-tailed test). We adjusted $P$-values to account for multiple comparisons using the Benjamini-Hochberg [@benjamini_controlling_1995] false discovery rate procedure implemented in the *R* package **multtest** version `r packageVersion("multtest")` [@pollard_multiple_2005].

For each sample image, we also simulated $10^3$ synthetic data with $n$ stomata ideally dispersed in an equilateral triangular grid. For these grids, we integrated over plausible stomatal densities and then conditioned on stomatal grids with exactly $n$ stomata. The simulated stomatal count was drawn from a Poisson distribution with the mean parameter $\lambda$ drawn from a Gamma distribution with shape $n$ and scale 1 $\lambda \sim \Gamma(n, 1)$. $\Gamma(n, 1)$ is the posterior distribution of $\lambda$ with a flat prior distribution. This allows us to integrate over uncertainty in the stomatal density from the sample image.

We developed a dispersion index $\mathrm{DI}$ to quantify how close observed stomatal distributions are to random uniform versus maximally dispered in an equilateral triangular grid. $\mathrm{DI}$ varies from zero to one, where zero is uniformly random and one is ideally dispersed:

\begin{equation}\label{eq:disp}
  \mathrm{DI} = \frac{\mathrm{NNI} - \text{median}(\mathrm{NNI_{random}})}{\text{median}(\mathrm{NNI_{uniform}}) - \text{median}(\mathrm{NNI_{random}})}
\end{equation}

\noindent $\mathrm{NNI}$ is calculated for each sample image as described above; $\text{median}(\mathrm{NNI_{random}})$ and $\text{median}(\mathrm{NNI_{uniform}})$ are calculated from the synthetic data specific to each sample image as described above. We tested whether light treatment affects $\mathrm{DI}$ and stomatla density ($D_S$) using analysis of variance (ANOVA).

<!-- CDM: I am not sure this part is needed
Uniformly distributed stomata on a leaf with stomatal density ($D_S$) maximize $\overline{D}_O$ in an equilateral triangle pattern with side lengths ($s$) equal to two times the incircle radius or apothem, $a$, of a regular hexagon with area ($A_\text{hex}$) equal to $A_\text{leaf}$ divided by the number of stomata ($n_\text{stomata}$) on the leaf according to Eq. \ref{eq:eq2}:

\begin{equation}\label{eq:eq2}
  s = \frac{\sqrt{2}} {3^{1/4} \sqrt{n_\text{stomata} / A_\text{leaf}}}
\end{equation}

As all sides of an equilateral triangle have the same length, the average nearest neighbor distance for a given $D_S$ is maximized and each stomate occupies the same hexagonal area ($A_\text{hex}$). As $DI$ approaches 1, the more uniform an area each of its stomata supply with CO$_2$ during photosynthesis. 
 -->

Finally, we examined the relationship between stomatal zone area and stomatal length using a Bayesian linear mixed-effects model fit with the *R* package **brms** version `r fit$version$brms` [@burkner_brms_2017; @burkner_advanced_2018] and *Stan* version `r fit$version$cmdstan` [@stan_development_team_stan_2023]. Stomatal zone area was calculated using Voronoi tessellation (e.g. \autoref{fig:tessellation-example}). The stomatal zone area, $S_\text{area}$, is the region of the leaf surface whose distance to stomate, $S$, is less than the distance to any other stomate, $S$. Stomatal length was measured in ImageJ [@schneider_nih_2012].

\begin{figure}[ht]
\includegraphics[width = \textwidth]{figures/tessellation-example.pdf}
\caption{Examples of synthetic and real leaf surfaces.  A) Uniform random synthetic leaf surface; B) Example of real leaf surface; C) Regularly distributed synthetic leaf surface. The zone defined by each stomate was calculated with voronoi tessellation and correlated with stomatal length in real leaves.}
\label{fig:tessellation-example}
\end{figure}

## Paired Abaxial and Adaxial Surface Analysis

To test whether the position of ab- and adaxial stomata are coordinated we compared the observed distribution to a null distribution where the positions on each surface are random. For each pair of surfaces (observed or synthetic) we calculated the distance squared between each pixel to the nearest stomatal centroid with the *R* package **raster** version `r packageVersion("raster")`. We refer to this as the 'nearest stomatal distance' or NSD. Then we calculated the pixel-wise Pearson correlation coefficient. If stomatal positions on each surface are coordinated to minimize the distance between mesophyll and the nearest stomate, then we expect a negative correlation. A pixel that is far from a stomate on one surface should be near a stomate on the other surface (\autoref{fig:ideal-amphi-grid}). We generated a null distribution of the correlation coefficient by simulating $10^3$ synthetic data sets for each observed pair. For each synthetic data set, we simulated stomatal position using a random uniform distribution, as described above, matching the number of stomata on abaxial and adaxial leaf surfaces. Stomatal positions on each surface are coordinated if the correlation coefficient is greater than 95% of the synthetic correlation values (one-tailed test).

## Modeling Photosynthesis

We modeled photosynthesis CO$_2$ assimilation rate using a spatially-explicit two-dimensional reaction diffusion model using a porous medium approximation [@parkhurst_intercellular_1994] using the finite element method (FEM) following @earles_excess_2017. Consider a two-dimensional leaf where stomata occur on each surface in a regular sequence with interstomatal distance $U$. The main outcome we assessed is the advantage of offsetting the position of stomata on each surface compared to have stomata on the same $x$ position on each surface. With these assumptions, by symmetry, we only need to model two stomata, one abaxial and one adaxial, from $x = 0$ to $x = U/2$ and from the adaxial surface at $y = 0$ to the abaxial surface at $y = L$, the leaf thickness. We arbitrarily set the adaxial stomate at $x = 0$ and toggled the abaxial stomata position between $x = U/2$ (offset) or $x = 0$ (below adaxial stomate). The 'coordination advantage' of offset stomatal position on each surface is the photosynthetic rate of the leaf with offset stomata compared to that with stomata aligned in the same $x$ position:

\begin{equation} \label{eq:coordination_advantage}
  \text{coordination advantage} = \frac{A_\text{offset}}{A_\text{aligned}}
\end{equation}

We modeled the coordination advantage over a range of leaf thicknesses, stomatal densities, photosynthetic capacities, and light environments to understand when offsetting stomatal position on each surface might deliver a significant photosynthetic advantage (\autoref{tab:model_var}). The complete model description is available in the Supporting Information.

```{r, model_var, results="asis"}
#| label: tbl-model_var

symbol_descriptions = left_join(rename(model_var, symbol = Variable),
                                pm_pars,
                                by = join_by(symbol)) |>
  transmute(x = glue("{symbol}: {name}")) |>
  pull(x) |>
  str_c(collapse = "; ")

# print(xtable::xtable(model_var, caption = paste0("The parameter range of model variables tested for their effect on coordination advantage (\\autoref{eq:coordination_advantage}) using a 2-D porous medium approximation. We used regularly spaced values within each range and simulated across all combinations. Here we converted model units to more conventional units (e.g. m to $\\mu$m). ", symbol_descriptions), label = "tab:model_var"),
#       comment = FALSE, include.rownames = FALSE, type = "latex")

model_var_caption = paste0("The parameter range of model variables tested for their effect on coordination advantage (\\autoref{eq:coordination_advantage}) using a 2-D porous medium approximation. We used regularly spaced values within each range and simulated across all combinations. Here we converted model units to more conventional units (e.g. m to $\\mu$m). ", symbol_descriptions)

knitr::kable(
  model_var, 
  caption = model_var_caption,
  label = "model_var",
  booktabs = TRUE, format = "latex", escape = FALSE
)

```

# Results

Stomatal density of *Arabidopsis thaliana* the `r nrow(single_surface_results)` leaf surfaces measured range from 12 to 93 (units) with high light leaves ranging from 93 to 55 (units), medium light from 15 to 35 (units), and low light from 12 to 42 (units). Stomatal density varies among light treatments (ANOVA, $`r report_fstat(density_anova, 'light')`$) because the density is much greater in the high light treatment (\autoref{fig:density}). Density is consistently greater on abaxial leaf surfaces across all light treatments (ANOVA, $`r report_fstat(density_anova, 'surface')`$; \autoref{fig:density}). There is no evidence for an interaction between light treatment and surface (ANOVA, $`r report_fstat(density_anova, 'light:surface')`$). Leaves are amphistomatous with a mean stomatal density ratio of 0.45.

\begin{figure}[ht]
\includegraphics[width = \textwidth]{figures/density.pdf}
\caption{Stomatal density is higher in \textit{A. thaliana} plants grown under high light conditions. We determined statistical significance between light treatments using Tukey post-hoc tests. $^*~0.05 > P \ge 0.01; ^{**}~0.01 > P \ge 0.001; ^{***}~0.0001 > P \ge 0.0001; ^{***}~ P <0.0001$.}
\label{fig:density}
\end{figure}

## Stomatal distribution is nonrandom, but far from ideal

Many leaf surfaces (`r nni2[['n1']]` of `r nni2[['n1']] + nni2[['n2']]`, `r round(100 * nni2[['n1']] / (nni2[['n1']] + nni2[['n2']]), 1)`%) are significantly overdispersed compared to a random uniform distribution, but none were close to an ideal hexagonal pattern (dispersion index = 1; \autoref{fig:single-surface}). Before controlling for multiple comparisons, `r round(100 * nni1[['n1']] / (nni1[['n1']] + nni1[['n2']]), 1)`% are significantly overdispersed. The dispersion index differs significantly among light treatments (ANOVA, $`r report_fstat(single_surface_anova, 'light')`$) because the medium light treatment is significantly less than the low treatment (\autoref{fig:single-surface}). Dispersion index is consistently greater on adaxial leaf surfaces across all light treatments (ANOVA, $`r report_fstat(single_surface_anova, 'surface')`$; \autoref{fig:single-surface}). There is no evidence for an interaction between light treatment and surface (ANOVA, $`r report_fstat(single_surface_anova, 'light:surface')`$).

\begin{figure}[ht]
\includegraphics[width=\textwidth]{figures/single-surface.pdf}
\caption{Stomata are more dispersed than expected under the null model of uniform random position (dispersion index = 0) but far from a distribution that maximizes distance between stomata (dispersion index = 1). We determined statistical significance between light treatments using Tukey post-hoc tests. $^*~0.05 > P \ge 0.01; ^{**}~0.01 > P \ge 0.001; ^{***}~0.0001 > P \ge 0.0001; ^{***}~ P <0.0001$.}
\label{fig:single-surface}
\end{figure}

## No evidence for coordinated stomatal position between surfaces

There is no evidence of spatial coordination between abaxial and adaxial leaf surfaces. The pixel-wise correlation between nearest stomatal distance (NSD) squared on paired abaxial and adaxial leaf surfaces is not significantly less than zero in any of the `r nrow(dual_surface_results)` leaves (\autoref{fig:dual-surface}). Before controlling for multiple comparisons, `r round(100 * nsd1[['n1']] / (nsd1[['n1']] + nsd1[['n2']]), 1)`% are significantly  *positively* correlated. The NSD correlation is not different among light treatments (ANOVA, $`r report_fstat(correlation_anova, 'light')`$; \autoref{fig:dual-surface}).

\begin{figure}[ht]
\includegraphics[width = \textwidth]{figures/dual-surface.pdf}
\caption{Pixel-wise correlation between near stomatal distance (NSD) squared on paired abaxial and adaxial leaf surfaces. Dashed line indicates zero correlation. Weak positive correlations are not significantly different from zero after correcting for multiple comparisons. The correlation does not differ among light treatments.}
\label{fig:dual-surface}
\end{figure}

## Larger stomata supply larger mesophyll volumes

Across all light treatments and leaf surfaces, stomatal length and stomatal area are weakly positively correlated, indicating some support for our hypothesis that stomata are selected for an optimal operational aperture (\autoref{fig:length-area}).

\begin{figure}[ht]
\includegraphics[width = \textwidth]{figures/length-area.pdf}
\caption{Stomatal length and stomatal zone area are positively correlated. Linear regression lines and 95\% confidence ribbons are from a Bayesian linear mixed-effects model.}
\label{fig:length-area}
\end{figure}

# Discussion

Stomata are expensive. A theoretical, optimized plant would minimize stomatal density while also allowing competitive gas exchange rates for its environment so as to maximize C assimilation per unit investment in stomata. Natural selection operates within developmental and physical constraints to drive each plant species toward its theoretical optimum. This study provides evidence that stomata in *Arbidopsis thaliana*, the model angiosperm, are non-randomly distributed, favoring dispersion over clustering (\autoref{fig:single-surface}). However, stomata are not ideally dispersed in an equilateral triangular grid as would be optimal to minimize CO$_2$ diffusion path length and standardize the area supplied by each stomate (\autoref{fig:tessellation-example}). Additionally, when grown in high light environments, *A. thaliana* exhibited increased stomatal density rather than increased stomatal dispersion (\autoref{fig:density}), which suggests that natural selection has acted more strongly on developmental pathways that modulate stomatal density than those that control stomatal dispersion. In other words, plants optimize gas exchange by adding more stomata rather than dispersing them more evenly across the leaf surface. This study also demonstrates that stomata that supply larger leaf areas with CO$_2$ tend to be larger (\autoref{fig:length-area}). These results could suggest that 1) the added energetic and hydraulic cost of non-ideally dispersed stomata is negligible and therefore not acted on by natural selection; 2) no developmental pathway exists to ensure the ideal placement of stomata on the leaf; or 3) the regulation of stomatal size limits the cost incurred by non-ideal stomatal dispersion.

In high light environments, amphistomy is favorable as high light photosynthesis is limited by CO$_2$ and amphistomy halves diffusion path length and boundary layer resistance, thereby reducing CO$_2$ limitation - increasing theoretical $A_\text{max}$. An optimal amphistomatous leaf has offset stomata such that stomata are more likely to appear on one leaf surface if there is not a stomata directly opposite it on the other surface as shown in \autoref{fig:ideal-amphi-grid}. However, our results show that leaf surfaces are not coordinated but are independent, regardless of light (\autoref{fig:dual-surface}). Additionally, gas exchange models show little photosynthetic efficiency gain from abaxial-adaxial stomatal coordination compared to anticoordination (INSERT FIG FROM MODELING). We posit that this marginal gain is not sufficient to be acted upon strongly by natural selection. Thus, amphistomatous plants do not exhibit abaxial-adaxial stomatal coordination for there is little selective advantage of it.

Our study corroborates previous studies which demonstrate that stomata are non-randomly distributed along the leaf surface as a result of developmental mechanisms such as spatially biased arrest of stomatal initials [@boetsch_arrest_1995], oriented asymmetric cell division [@geisler_oriented_2000], and cell cycle controls [@croxdale_stomatal_2000]. We do not investigate the potential developmental pathways that influence stomatal dispersion in this study; however, they are important to consider as these pathways could limit plants from reaching the theoretical peak in the adaptive landscape: uniform stomatal dispersion. Instead, as this study suggests, plants may simply compensate with higher stomatal density and by fitting stomatal size to the area that they supply with CO$_2$. To understand why stomata are not ideally dispersed, more modelling should be done to estimate the fitness gain of stomatal dispersion. Additionally, genetic manipulation studies should attempt to create mutants with clustered and ideally dispersed stomata for a comparison of their photosynthetic traits. This could have extremely important implications for maximum assimilation rates in crops as most crop species are grown in high light where CO$_2$ is often limiting. In drought-prone environments, increased stomatal dispersion may increase water use efficiency by reducing the number of stomata needed to achieve the same internal CO$_2$ concentration, $C_\text{i}$.

Beyond disperson on a single surface, gas exchange can be optimized via stomatal coordination of abaxial and adaxial surfaces in amphistomatous leaves. Given that leaf thicknesses are generally multiple times greater than interstomatal distance (GIVE DISTANCES HERE). As a result, abaxial-adaxial stomatal coordination reduces CO$_2$ diffusion path length far less than single surface dispersion, so we hypothesize this strategy to afford less photosynthetic advantage to the leaf. Our modelling results demonstrate that, even in ideal conditions, i.e. thick leaf, low stomatal densities, high light, low leaf porosity, high Rubisco concentration, etc., the photosynthetic advantage of coordination is minimal. We are not surprised by these results, but still highlight them here as we are the first to report this finding.

Amphistomy is a unique and important adaptation found around the world across many plant lineages [@muir_light_2018], yet much of the dynamics of amphistomy remain poorly understood. Here, we show that in *Arabidopsis thaliana* 1) stomata in are non-randomly dispersed, but not ideally dispersed; 2) stomatal size and density are modulated by light; 3) stomatal size is positively correlated with the area to which it supplies CO$_2$; and 4) abaxial-adaxial stomatal coordination is not exhibited and is not shown to provide a strong photosynthetic advantage using CO$_2$ diffusion models. Interestingly, these findings did not validate many of our hypotheses which were based on first principles, suggesting that there may be limits on plants' ability to control stomatal placement. Future studies which elucidate these limitations may have important implications for agriculturaly productivity in a rapidly changing world.

# References